Inferring the exoplanet radius distribution from a noisy, incomplete catalog

### 2013-11-26 document/ms.tex
new title

\title{%
    Exoplanet population inferences
    from a hierarchical model
    of the \kepler\ ``Object of Interest'' Catalog
}

@@

\section{A toy model}

We start by assuming that ever target star has some $K$ orbiting planets.
Most of these planets are unobserved because they are two (2) small to be
detectable or because of geometry.
The graphical model in \fig{gm} is a sketch of the hypothetical generative
procedure for a catalog of exoplanet observations.
The way it is shown, this model is extremely general but we will consider a
very specific form here.
The parameters of this model are listed and described in \tab{parameters}.

Using this model, we want to find---conditioned on the catalog of exoplanet
candidates---joint constraints on the parameters describing the population of
exoplanets and the selection function of the survey.
To this end, we need to be able to efficiently marginalize over the parameters
of all of the exoplanet systems (observed and unobserved) to compute the
marginalized likelihood function
\begin{eqnarray}
p(\planetobs,\stellarobs,\isobs\,|\,\population,\selection) &=&
\prod_{n=1}^N \int \left [
p(\stellar_n)\,p(\stellarobs_n\,|\,\stellar_n) \,
 \prod_{k=1}^K
p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection)\right ]
\dd \stellar_n
\end{eqnarray}
where
\begin{eqnarray}\eqlabel{planet-marg}
p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
\int
p(\planet_{kn}\,|\,\population) \,
p(\planetobs_{kn}\,|\,\planet_{kn},\stellar_{n}) \,
p(\isobs_{kn}\,|\,\planet_{kn},\stellar_{n},\selection)
\dd \planet_{kn} \quad.
\end{eqnarray}

We can evaluate the integral in \eq{planet-marg} using importance sampling.
If the exoplanet is not observed ($\isobs_{kn}=0$), the most sensible proposal
density is simply $\tilde{p}(\planet_{kn})=p(\planet_{kn}\,|\,\population)$
and the integral becomes
\begin{eqnarray}
p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
\frac{1}{\sum_l \tilde{p}(x_l)}\,\sum_l
p(\isobs_{kn}\,|\,\planet_{kn}=x_l,\stellar_{n},\selection)
\end{eqnarray}
for $x_l \sim p(\planet_{kn}\,|\,\population)$.

@@

\begin{deluxetable}{ccl}
\tablecaption{%
A description of the model parameters.
\tablabel{parameters}}
\tablewidth{0pt}
\tablehead{& & \colhead{Description}}
\startdata
\multirow{3}{*}{\population}
& $\population_\period$ & The period distribution \\
& $\population_\radius$ & The exoplanet radius distribution \\
& $\population_\relincl$ & The relative inclination distribution \\
\hline
\selection & --- & The parameters of the catalog selection function \\
\hline
$\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
star $n$ in the catalog \\
\hline
\multirow{3}{*}{$\planet_{kn}$}
& $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
& $\radius_{kn}$ & The exoplanet's radius \\
& $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
\hline
\multirow{3}{*}{$\planetobs_{kn}$}
& $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
& $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
& $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
\hline
\multirow{4}{*}{$\stellar_{n}$}
& $\smass_{n}$ & The mass of star $n$ \\
& $\sradius_{n}$ & The radius of star $n$ \\
& $\snoise_{n}$ & The ``variability'' of star $n$ \\
& $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
\hline
\multirow{2}{*}{$\stellarobs_{n}$}
& $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
& $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
\enddata
\end{deluxetable}

>>>>

\title{%
Inferring the exoplanet radius distribution from a noisy, incomplete catalog
}

@@

@@

% \begin{deluxetable}{ccl}
% \tablecaption{%
% A description of the model parameters.
% \tablabel{parameters}}
% \tablewidth{0pt}
% \tablehead{& & \colhead{Description}}
% \startdata
% \multirow{3}{*}{\population}
% & $\population_\period$ & The period distribution \\
% & $\population_\radius$ & The exoplanet radius distribution \\
% & $\population_\relincl$ & The relative inclination distribution \\
% \hline
% \selection & --- & The parameters of the catalog selection function \\
% \hline
% $\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
% star $n$ in the catalog \\
% \hline
% \multirow{3}{*}{$\planet_{kn}$}
% & $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
% & $\radius_{kn}$ & The exoplanet's radius \\
% & $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
% \hline
% \multirow{3}{*}{$\planetobs_{kn}$}
% & $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
% & $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
% & $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
% \hline
% \multirow{4}{*}{$\stellar_{n}$}
% & $\smass_{n}$ & The mass of star $n$ \\
% & $\sradius_{n}$ & The radius of star $n$ \\
% & $\snoise_{n}$ & The ``variability'' of star $n$ \\
% & $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
% \hline
% \multirow{2}{*}{$\stellarobs_{n}$}
% & $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
% & $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
% \enddata
% \end{deluxetable}
